Variation in gaze understanding across the life span: A process-level perspective

### 2023-06-09 paper/manuscript.tex
add gazemodel data adults; write magnet sections

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{social cognition, individual differences, gaze cues, cognitive development\newline\indent Word count: 6856}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{setspace}
\captionsetup[figure]{font={stretch=1}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={TANGO: A reliable, open-source, browser-based task to assess individual differences in gaze understanding in 3 to 5-year-old children and adults},
  pdfauthor={Julia Prein1, Manuel Bohn1, Steven Kalinke1, \& Daniel B. M. Haun1},
  pdflang={en-EN},
  pdfkeywords={social cognition, individual differences, gaze cues, cognitive development},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{TANGO: A reliable, open-source, browser-based task to assess individual differences in gaze understanding in 3 to 5-year-old children and adults}
\author{Julia Prein\textsuperscript{1}, Manuel Bohn\textsuperscript{1}, Steven Kalinke\textsuperscript{1}, \& Daniel B. M. Haun\textsuperscript{1}}
\date{}

\shorttitle{assessing individual differences in gaze understanding}

@@

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}

\abstract{%
Traditional measures of social cognition used in developmental research often lack satisfactory psychometric properties and are not designed to capture variation \emph{between} individuals.
Here we present TANGO (Task for Assessing iNdividual differences in Gaze understanding - Open); a brief (approx. 5--10min), reliable, open-source task to quantify individual differences in the understanding of gaze cues. Localizing the attentional focus of an agent is crucial in inferring their mental states, building common ground, and thus, supporting cooperation.
Our interactive browser-based task works across devices and enables in-person and remote testing. The implemented spatial layout allows for discrete and continuous measures of participants' click imprecision and is easily adaptable to different study requirements.
Our task measures inter-individual differences in a child (N = 387) and an adult (N = 236) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test--retest reliability estimates underline that the captured variation is systematic. This work shows a promising way forward in studying individual differences in social cognition and will help us explore the structure and development of our core social-cognitive processes in greater detail.
}

@@

Social cognition -- representing and reasoning about an agent's perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior -- is among the most-studied phenomena in developmental research.
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development (Gopnik \& Slaughter, 1991; Peterson, Wellman, \& Slaughter, 2012; Rakoczy, 2022; Wellman, Cross, \& Watson, 2001; Wellman \& Liu, 2004).
Yet, there are always individual differences.
Identifying variability in social-cognitive abilities and factors influencing their development is vital in theory building (e.g., to test causal predictions) and designing interventions (Happé, Cook, \& Bird, 2017; Kidd, Donnelly, \& Christiansen, 2018; Lecce, Bianco, Devine, Hughes, \& Banerjee, 2014; Mundy et al., 2007; Underwood, 1975).

Numerous studies have already examined individual differences in social cognition (for an overview, see Hughes \& Devine, 2015; Slaughter, 2015).
The most common, recurring research questions are concerned with the developmental sequence of social-cognitive abilities (e.g., Wellman \& Liu, 2004), and which factors drive the development of social cognition (Devine \& Hughes, 2018; e.g., Gola, 2012).
For example, Okumura and colleagues asked how early gaze-following and object processing relate to later language development (Okumura, Kanakogi, Kobayashi, \& Itakura, 2017).
In general, individual differences studies often focus on the relationship between social-cognitive abilities and: (1) family influences, (2) other cognitive constructs, and (3) social behavioral outcomes (for an overview, see Repacholi, 2003).
Studies on social-cognitive abilities and family influences include the effect of parenting practices (for a review, see Pavarini, de Hollanda Souza, \& Hawk, 2013), attachment quality (e.g., Astor et al., 2020), mental state talk (Gola, 2012; Hughes, Ensor, \& Marks, 2011; Lecce et al., 2014), and family background as parental education, occupation, sibling interaction and childcare (Bulgarelli \& Molina, 2016; Cutting \& Dunn, 1999; Dunn, Brown, Slomkowski, Tesla, \& Youngblade, 1991).
Another group of individual differences studies focuses on the interplay of social and physical cognition (Herrmann, Hernández-Lloreda, Call, Hare, \& Tomasello, 2010), executive functions (Benson, Sabbagh, Carlson, \& Zelazo, 2012; Buttelmann, Kühn, \& Zmyj, 2021; Carlson \& Moses, 2001; Carlson, Moses, \& Claxton, 2004; Hughes \& Ensor, 2007), and language abilities (McEwen et al., 2007; Milligan, Astington, \& Dack, 2007; Okumura et al., 2017).
Studies on social behavioral outcomes measured the interplay of social cognition and prosociality (for a review, see Imuta, Henry, Slaughter, Selcuk, \& Ruffman, 2016; Walker, 2005), stereotypes, and resource allocations (Rizzo \& Killen, 2018), and moral intentions (Sodian et al., 2016).

However, developmental psychologists are frequently surprised to find minor or no association between measures of social cognition that are thought to be theoretically related -- cross-sectionally and/or longitudinally (e.g., Poulin-Dubois, Goldman, Meltzer, \& Psaradellis, 2023; Sodian, 2023; Sodian et al., 2016).
This might be because traditional measures of social cognition are not designed to capture variation between children: they often rely on low trial numbers, small sample sizes, and dichotomous measures.
A recent review showed that many studies on social cognition measures failed to report relevant psychometric properties at all (Beaudoin, Leblanc, Gagner, \& Beauchamp, 2020) or -- when they did -- showed mixed results on test-retest reliability (Hughes et al., 2000; Mayes, Klin, Tercyak, Cicchetti, \& Cohen, 1996).

To give an example: the most commonly applied prototypical measure for social cognition is the change-of-location false belief task (Baron-Cohen, Leslie, \& Frith, 1985; Wimmer \& Perner, 1983).
Here, children watch a short sequence of events (often acted out or narrated by the experimenters).
A doll called Sally puts her marble into a basket.
After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box.
Participants then get asked where Sally will look for her marble once she returns.
The outcome measures false belief understanding in a dichotomous way: children pass the task if they take the protagonist's epistemic state into account and answer that she will look into the basket.
Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age (for a review, see Wellman et al., 2001).
Several cross-cultural studies supported this evidence (Barrett et al., 2013; Callaghan et al., 2005; cf. Mayer \& Träuble, 2015).

However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value (Repacholi, 2003).
Thus, this task seems well suited to track a particular group-level developmental transition, yet it fails to capture individual differences (cf.~``reliability paradox,'' Hedge, Powell, \& Sumner, 2018).
As Wellman (2012) put it, ``it's really only passing/failing one sort of understanding averaged across age'' (p.~317).
This has profound implications for what studies on individual differences using this task (or others) can show.
Poor measurement of social cognition on an individual level is likely to conceal relations between different aspects of cognition and may obscure developmental change.
For example, Sodian et al. (2016) neither found a correlation between two moral Theory of mind False Belief and Intention tasks at 60 months, nor a relationship between these two factors and implicit False Belief understanding at 18 months.

The ``Sandbox task'' is one of the few tasks that attempt to overcome these methodological challenges (Begeer, Bernstein, van Wijhe, Scheeren, \& Koot, 2012; Bernstein, Thornton, \& Sommerville, 2011; Coburn, Bernstein, \& Begeer, 2015; Mahy, Bernstein, Gerrard, \& Atance, 2017; Sommerville, Bernstein, \& Meltzoff, 2013).
This continuous FB task measures the degree to which the estimate of another's belief is biased by one's own knowledge.
Recent work questions the interpretation of this measure (Samuel, Legg, Lurz, \& Clayton, 2018a; Samuel, Legg, Lurz, \& Clayton, 2018b): it is unclear whether a smaller egocentric bias can be directly translated into a better mental state reasoning ability.
Another evaluation criterion should, therefore, be whether a task captures meaningful variability in performance; that is, differences in test scores should correspond to differences in the social-cognitive ability in question.

Thus, developmental psychology faces a dilemma: many research questions rely on measuring individuals' development, yet, there is a lack of tasks to measure these individual differences reliably.
To capture the emergence of social-cognitive abilities and their relation to social factors in greater precision and detail, we must consequently address the methodological limitations of existing study designs (Hughes et al., 2011; Hughes \& Leekam, 2004).

Schaafsma, Pfaff, Spunt, and Adolphs (2015) compiled a ``wish list'' for new social-cognitive paradigms.
They advocated for parametric -- instead of dichotomous -- measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates (see also Beaudoin et al., 2020; Hughes \& Devine, 2015).
New tasks should capture variation across age groups, including older children and adults (Repacholi, 2003).
Another goal in creating new tasks should be to focus on the ``face value'': measures should probe the underlying social-cognitive ability as straight-forward and directly as possible.
Keeping task demands minimal is also beneficial for using the paradigm in a variety of different cultural, clinical, and demographic contexts (Molleman, Kurvers, \& van den Bos, 2019).
The task should serve as a proxy for behavior as it appears in the real world and should be validated in relation to real-world experiences (Repacholi, 2003).

\hypertarget{a-new-measure-of-gaze-understanding}{%
\section{A new measure of gaze understanding}\label{a-new-measure-of-gaze-understanding}}

Our goal was to design a new measure of social cognition that captures individual differences across age groups in a systematic and reliable way.
We focused on a fundamental ability implicated in many social-cognitive reasoning processes: gaze understanding -- the ability to locate and use the attentional focus of an agent.
The first component of this ability is often termed gaze following -- turning one's eyes in the same direction as the gaze of another agent -- and has been studied intensively (Astor, Thiele, \& Gredebäck, 2021; Byers-Heinlein et al., 2021; Coelho, George, Conty, Hugueville, \& Tijus, 2006; Del Bianco, Falck-Ytter, Thorup, \& Gredebäck, 2019; Frischen, Bayliss, \& Tipper, 2007; Hernik \& Broesch, 2019; Itakura \& Tanaka, 1998; Lee, Eskritt, Symons, \& Muir, 1998; Moore, 2008; Shepherd, 2010; Tomasello, Hare, Lehmann, \& Call, 2007).
In our definition, gaze understanding goes one step further by including the \emph{acting on the gaze-cued location} -- therefore, using the available social information to guide one's behavior as needed in real-life conditions.

Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a ``front end ability'' (Brooks \& Meltzoff, 2005, p. 535).
Gaze is integral for many more sophisticated social-cognitive abilities, for example, inferences about knowledge states.
As such, the eyes have been regarded as a ``window into the mind'' (Shepherd, 2010).
Monitoring another's attention also supports building a common ground, which is important for action coordination and cooperative social interactions (Bohn \& Köymen, 2018; Tomasello et al., 2007).
In addition, gaze and language development seem to be related (Brooks \& Meltzoff, 2005).
Gaze facilitates word learning by helping to identify the referent of a new word and has been regarded as a crucial signal of nonverbal communication (Hernik \& Broesch, 2019; Macdonald \& Tatler, 2013).
While the emergence of gaze following has been well established, less is known about the developmental trajectory throughout childhood and adolescence.
One possibility is that our social-cognitive ability in question is fully developed once emerged in infancy.
However, many cognitive abilities continue to develop beyond early childhood (e.g., Gathercole, Pickering, Ambridge, \& Wearing, 2004 for working memory; Raviv \& Arnon, 2018 for visual statistical learning).
Therefore, children could potentially improve in understanding gaze, fine-tuning the performance of the already existing skill.
Consequently, we aimed to assess the differentiation of the ability to understand gaze.
Our goal was not to establish the youngest age at which children understand gaze cues.
Rather, we wanted to examine how that ability changes with age.
To accurately measure developmental change, we were interested in capturing individual variability.

To address the psychometric shortcoming of earlier work, we implemented the following design features: First, we used a continuous measure which allowed us to capture fine-grained individual differences at different ages.
Second, we designed short trials that facilitate more than a dozen replicates per subject.
The result is more precise individual-level estimates.
Third, we systematically investigated the psychometric properties of the new task.

Designing this task required a new testing infrastructure.
We designed the task as an interactive web application.
Previous research has successfully used online study implementations that compare well to in-person data collection (Bohn, Le, Peloquin, Köymen, \& Frank, 2021; Bohn, Tessler, Merrick, \& Frank, 2021; Frank, Sugarman, Horowitz, Lewis, \& Yurovsky, 2016).
This greatly increased the flexibility with which we could modify the stimuli on a trial-by-trial basis.
Furthermore, because the task is largely self-contained, it is much more controlled and standardized.
Most importantly, it makes the task portable: testing is possible in-person using tablets but also remotely via the internet (no installation needed).
As such, it provides a solid basis to study individual differences in gaze understanding across ages at scale.
We make the task and its source code openly accessible for other researchers to use and modify.

\hypertarget{task-design}{%
\section{Task design}\label{task-design}}

\hypertarget{implementation}{%
\subsection{Implementation}\label{implementation}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{social cognition, individual differences, gaze cues, cognitive modeling\newline\indent Word count: xxx}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{setspace}
\captionsetup[figure]{font={stretch=1}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Variation in gaze understanding across the life span: A process-level perspective},
  pdfauthor={Julia Prein1, Manuel Bohn1,2, Luke Maurits1, Annika Werwach1, \& Daniel B. M. Haun1},
  pdflang={en-EN},
  pdfkeywords={social cognition, individual differences, gaze cues, cognitive modeling},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Variation in gaze understanding across the life span: A process-level perspective}
\author{Julia Prein\textsuperscript{1}, Manuel Bohn\textsuperscript{1,2}, Luke Maurits\textsuperscript{1}, Annika Werwach\textsuperscript{1}, \& Daniel B. M. Haun\textsuperscript{1}}
\date{}

\shorttitle{modeling variation in gaze understanding}

@@

The authors made the following contributions. Julia Prein: Conceptualization, Methodology, Software, Investigation, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Manuel Bohn: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Luke Maurits: Formal Analysis, Writing - Review \& Editing; Annika Werwach: Methodology, Investigation, Writing - Review \& Editing; Daniel B. M. Haun: Supervision, Writing - Review \& Editing.

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany\\\textsuperscript{2} Institute of Psychology, Leuphana University Lüneburg, Germany}

\abstract{%
Abstract
}

@@

\begin{itemize}
\tightlist
\item
  why do we care about developmental trajectory? ref to stat learning paper
\item
  variation
\end{itemize}

\hypertarget{why-do-we-need-gaze-understanding}{%
\subsection{Why do we need gaze understanding?}\label{why-do-we-need-gaze-understanding}}

How do humans learn about their environment and navigate through their social surroundings?
One possibility to extract information from the environment is through following others' focus of attention.
Building a common ground is considered especially important in communicative interactions and shared activities (Tomasello, Hare, Lehmann, \& Call, 2007).

\hypertarget{how-does-gaze-following-emerge}{%
\subsection{How does gaze following emerge?}\label{how-does-gaze-following-emerge}}

Existing studies operationalize gaze following as the ability to follow another agent's line of sight.
As one of the most fundamental social-cognitive abilities, it has been extensively studied in infancy and early childhood.
Infants as young as six months can attune their gaze to that of another agent (D'Entremont, Hains, \& Muir, 1997).
At the end of their first year of life, infants can follow gaze to locations outside their current visual field and move themselves to gain proper perceptual access (Moll \& Tomasello, 2004).

While the emergence of gaze following has been well established, less is known about the developmental trajectory throughout childhood and adolescence.
One possibility is that our social-cognitive ability in question is fully developed once emerged in infancy.However, many cognitive abilities develop with age (e.g., working memory, Gathercole, Pickering, Ambridge, \& Wearing, 2004).
Similarly, visual processing appears to improve with age.
Therefore, children could potentially improve in gaze following, fine-tuning the performance of the already existing skill.

\hypertarget{the-scope-of-infants-gaze-following-ability}{%
\subsection{The scope of infants' gaze following ability}\label{the-scope-of-infants-gaze-following-ability}}

Though these studies suggest that young infants can align their visual attention to another's line of sight, it does not necessarily include understanding the intentions of the other agent.
Infants could simply attune their orientation or be attracted by others' gaze without processing what exactly the other is seeing (cf.~Butterworth \& Jarrett's ecological and geometric mechanism, Butterworth and Jarrett (1991){]}. Therefore, it is crucial to study children's intentional understanding of gaze.

Moore, Angelopoulos, and Bennett (1997) showed that 9-month-olds followed an agent's gaze more, when it was accompanied by a dynamic head turn in comparison to a static head turn.

In a hiding game with two search locations, Povinelli, Reaux, Bierschwale, Allain, and Simon (1997) found that three-year-olds used gaze as a cue to locate the reward, while two-year-olds performed at chance level.

In a similar object choice paradigm with two containers, Behne, Carpenter, and Tomasello (2005) investigated whether infants understand the communicative intent behind pointing and gaze cues.
In contrast to Povinelli et al. (1997), they found that already 14-month-olds used the agent's cues to select an object.
In conditions with absent-minded `cues', infants performed around chance.
This could be interpreted as infants recognizing the nature of this joint activity: namely, that the adult's behavior was beneficial and relevant for their object choice.

\hypertarget{head-vs-eye-direction}{%
\subsubsection{Head vs eye direction}\label{head-vs-eye-direction}}

It is important to note that in many existing gaze conditions, the experimenter shifted their eyes and head in synchrony (e.g., Behne et al. (2005)).
Instead of pointing towards gaze understanding, a critic could claim that the results can be explained by face direction alone.

A handful of studies approached this potential confound by separately manipulating head and eye movement.
Brooks and Meltzoff (2002) implemented a comparison between eye and head orientation and found that 14-month-olds were sensitive to open versus closed eyes.

Investigating the `cooperative eye hypothesis', Tomasello et al. (2007) implemented six conditions, in which an experimenter oriented towards the ceiling with their eyes only, head only (eyes closed), both head and eyes, or neither.
They found that human infants relied more on the eye movement, while chimpanzees paid more attention to the head movement.

Importantly, the subjects were not presented with an object choice but their attention orientation was measured.

\begin{itemize}
\item
  (Raviv \& Arnon, 2018)
\item
  (Astor \& Gredebäck, 2022)
\item
  (Colombo, 2001)
\item
  (Scaife \& Bruner, 1975)
\item
  (Itakura \& Tanaka, 1998)
\item
  (Carpenter, Nagell, \& Tomasello, 1998) ``Several other studies have attempted to determine more precisely the cue that infants are using when they follow the gaze direction of others, that is, whether they use adults' head or eye orientation. In tasks comparing infants' responses when the experimenters turned their head and eyes together to targets with their responses when the experimenters directed their eyes to the targets but their head remained facing forward, Corkum and Moore (1995), Lempers (1979), and Lempers, Flavell, and Flavell (1977) all found that only infants age 12 months and older responded correctly when eyes and head were oriented in the same direction and that infants at all ages (i.e., through 19 months) performed poorly when eye and head direction diverged'' (p.10-11) object choice.
\item
  (Silverstein, Feng, Westermann, Parise, \& Twomey, 2021) for vertical plane
\item
  (Zhang, Zhang, Zhang, Tang, \& Liu, 2019)
\item
  (Frischen, Bayliss, \& Tipper, 2007)
\item
  (Lee, Eskritt, Symons, \& Muir, 1998)
\item
  (Coelho, George, Conty, Hugueville, \& Tijus, 2006)
\end{itemize}

\hypertarget{aim-of-the-current-project}{%
\subsection{Aim of the current project}\label{aim-of-the-current-project}}

\hypertarget{developmental-trajectory-measuring-modeling-individual-differences}{%
\subsubsection{Developmental trajectory, measuring \& modeling individual differences}\label{developmental-trajectory-measuring-modeling-individual-differences}}

In this study, we were interested in the developmental trajectory of gaze understanding.
While we expect the younger children to be able to follow gaze, we aimed at assessing the differentiation of their social-cognitive ability.
Our goal was \emph{not} to establish the youngest age at which children understand gaze cues.
Rather, we wanted to examine how that ability changes with age.

In our study, we focused on the communicative intents of gaze: we asked children to locate a target by following an agent's gaze.
While language demands were kept low, the participants had to actively respond and, therefore, make use of the presented gaze cue.

A unique contribution of this study is the richness of the data set.
Methodological challenges arise when trying to compare data across ages from qualitatively and quantitatively different study tasks.
We could circumvent these issues by applying the exact same task for the entire life span.

\hypertarget{lifespan}{%
\section{Lifespan}\label{lifespan}}

\begin{itemize}
\tightlist
\item
  development \& individual differences in gaze understanding
\item
  verweis methods paper reliable differences kinder \& adults.
\item
  kontinuerliche, systematische variation, wodrin? =\textgreater{} model
\end{itemize}

\url{https://osf.io/6yjz3}

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:lifespan_sample}}

\begin{tabular}{lllll}
\toprule
Age group & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Age mean} & \multicolumn{1}{c}{Age range} & \multicolumn{1}{c}{Age SD}\\
\midrule
3.00 & 19 (7 female) & 3.62 & 3.04 - 3.99 & 0.31\\
4.00 & 17 (9 female) & 4.45 & 4.05 - 4.91 & 0.30\\
5.00 & 22 (13 female) & 5.56 & 5.08 - 5.99 & 0.31\\
6.00 & 24 (16 female) & 6.50 & 6.1 - 6.99 & 0.28\\
7.00 & 39 (20 female) & 7.48 & 7.04 - 7.95 & 0.25\\
8.00 & 41 (20 female) & 8.46 & 8.03 - 8.98 & 0.27\\
9.00 & 56 (29 female) & 9.46 & 9.01 - 9.96 & 0.28\\
10.00 & 35 (22 female) & 10.49 & 10.01 - 11 & 0.28\\
11.00 & 54 (26 female) & 11.43 & 11.01 - 11.96 & 0.28\\
12.00 & 43 (19 female) & 12.41 & 12.01 - 12.99 & 0.30\\
13.00 & 42 (19 female) & 13.50 & 13.09 - 13.99 & 0.27\\
14.00 & 20 (14 female) & 14.37 & 14.05 - 14.98 & 0.23\\
15.00 & 21 (11 female) & 15.56 & 15.05 - 15.98 & 0.30\\
16.00 & 19 (10 female) & 16.51 & 16.17 - 16.97 & 0.24\\
17.00 & 19 (10 female) & 17.53 & 17.01 - 17.95 & 0.28\\
18.00 & 2 (0 female) & 18.00 & 18 - 18 & 0.00\\
19.00 & 5 (4 female) & 19.00 & 19 - 19 & 0.00\\
20.00 & 40 (25 female) & 23.02 & 20 - 29 & 2.77\\
30.00 & 40 (21 female) & 34.42 & 30 - 39 & 3.00\\
40.00 & 40 (24 female) & 44.17 & 40 - 49 & 2.92\\
50.00 & 40 (21 female) & 54.38 & 50 - 59 & 3.04\\
60.00 & 40 (21 female) & 63.73 & 60 - 69 & 2.56\\
70.00 & 40 (20 female) & 72.75 & 70 - 79 & 2.44\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

We collected data from a remote child, teenager and adult sample.
For the remote child and teenager sample, we recruited participants via an internal database consisting of families living in Leipzig, Germany, who volunteered to participate in child development studies and indicated an interest in online studies.

The remote child and teenager sample consisted of 471 children, including 19 3-year-olds (mean = 3.62 years, SD = 0.31, range = 3.04 - 3.99, 7 girls), 17 4-year-olds (mean = 4.45 years, SD = 0.30, range = 4.05 - 4.91, 9 girls), 22 5-year-olds (mean = 5.56 years, SD = 0.31, range = 5.08 - 5.99, 13 girls), 24 6-year-olds (mean = 6.50 years, SD = 0.28, range = 6.10 - 6.99, 16 girls), 39 7-year-olds (mean = 7.48 years, SD = 0.25, range = 7.04 - 7.95, 20 girls), 41 8-year-olds (mean = 8.46 years, SD = 0.27, range = 8.03 - 8.98, 20 girls), 56 9-year-olds (mean = 9.46 years, SD = 0.28, range = 9.01 - 9.96, 29 girls), 35 10-year-olds (mean = 10.49 years, SD = 0.28, range = 10.01 - 11, 22 girls), 54 11-year-olds (mean = 11.43 years, SD = 0.28, range = 11.01 - 11.96, 26 girls), 43 12-year-olds (mean = 12.41 years, SD = 0.30, range = 12.01 - 12.99, 19 girls), 42 13-year-olds (mean = 13.50 years, SD = 0.27, range = 13.09 - 13.99, 19 girls), 20 14-year-olds (mean = 14.37 years, SD = 0.23, range = 14.05 - 14.98, 14 girls), 21 15-year-olds (mean = 15.56 years, SD = 0.30, range = 15.05 - 15.98, 11 girls), 19 16-year-olds (mean = 16.51 years, SD = 0.24, range = 16.17 - 16.97, 10 girls), 19 17-year-olds (mean = 17.53 years, SD = 0.28, range = 17.01 - 17.95, 10 girls), 2 17-year-olds (mean = 18 years, SD = 0, range = 18 - 18, 0 girls),
5 17-year-olds (mean = 19 years, SD = 0, range = 19 - 19, 4 girls).
